ICS_Cytokine_Analysis

### 2019-06-17 Cytokine_Analysis.Rmd
narrowed down to LTBI and TB

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
```

@@

CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))
CD4b<-plot_filter(read.csv("ICS_Cytokine_new.csv"))

@@

Compass<-plot_filter(read.csv("ICS_Compass_scores_complete.csv"))

@@

data<-read.csv("ICS_Donor_Info_clean.csv")
data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
data$TB<-factor(data$TB, levels=c("Naïve","Healthy Control","LTBI","Active TB"))
data$SM<-gsub("US|Not US|Unknown", "Naïve", data$SM)
data$SM<-factor(data$SM, levels=c("Naïve","Negative","Schisto+"))
Donor<-data
```

@@

##Whole Cell Lysate
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="WCL") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_4(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
colnames<-c("TH1_total", "TH2_total")
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4b, CD4b$Stim=="WCL") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_4(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

##CFP-10 & ESAT-6
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_4(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
colnames<-c("TH1_total", "TH2_total")
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4b, CD4b$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_4(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

#Compass Graphs
```{r}
#CD4 Mtb Responses as measured by Compass
DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype == "CD4")

@@

```{r, fig.width=16, fig.height=12}
library(dplyr)
DF<-merge(Compass, dplyr::select(Donor, -X, -TB, -SM), by="Donor")
ggpairs(dplyr::filter(DF, celltype=="CD4", Stim=="WCL", TB!="N")%>%
            mutate(disease = paste(TB, SM, sep=" "))%>%
            dplyr::select(age, sex, FS, PFS, disease),
        aes(col=disease))

```

#Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.
```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
triple_boolean_plot(CD4, "PMA")
triple_boolean_plot(CD4, "WCL")
triple_boolean_plot(CD4, "PEP")
```

#T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups. 
```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
colnames<-c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq")
### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "WCL", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="WCL",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### Peptide
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PEP", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PEP",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```
\newpage

#Individual Cytokines
These are the frequencies of cells producing individual cytokines without considering whether or not they are single positive, double positive etc. So any cell producing IL-4 is included in the IL-4 analysis.

@@

```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)

colnames<-c("TH1_Total" , "TH2_Total" )

### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(CD4b, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             title="PMA & Ionomycin",
             x="",
             y=paste(col, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
```

@@

CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))

@@

Compass<-plot_filter(read.csv("ICS_Compass_scores_complete_clean.csv"))

@@

data<-read.csv("ICS_DonorInfo_clean.csv")
data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
data$TB<-factor(data$TB, levels=c("Naïve","Healthy Control","LTBI","Active TB"))
data$SM<-gsub("US|Not US|Unknown", "Naïve", data$SM)
data$SM<-factor(data$SM, levels=c("Naïve","Negative","Schisto+"))
Donor<-data
```

@@

##Whole Cell Lysate
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="WCL") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_4(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

##CFP-10 & ESAT-6
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_4(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

```{r}
grid.arrange(
    plot_2(subset(df, df$Stim=="PMA"), "disease", "IFNg")+
        labs(y="IFNg+ CD4+ T cells (%)", title="PMA & Ionomycin"),
    plot_2(subset(df, df$Stim=="PEP"), "disease", "IFNg")+
        labs(y="", title="CFP-10 & ESAT-6"),
    ncol=2)

grid.arrange(
    plot_2(subset(df, df$Stim=="PMA"), "disease", "G_T")+
        labs(y="IFNg+TNFa+ CD4+ T cells (%)", title="PMA & Ionomycin"),
    plot_2(subset(df, df$Stim=="PEP"), "disease", "G_T")+
        labs(y="", title="CFP-10 & ESAT-6"),
    ncol=2)

grid.arrange(
    plot_2(subset(df, df$Stim=="PMA"), "disease", "x_4_13_T")+
        labs(y="IL4+TNFa+ CD4+ T cells (%)", title="PMA & Ionomycin"),
    plot_2(subset(df, df$Stim=="PEP"), "disease", "x_4_13_T")+
        labs(y="", title="CFP-10 & ESAT-6"),
    ncol=2) 
```

#Extra Graphs

##Compass Graphs
```{r}
#CD4 Mtb Responses as measured by Compass
DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype == "CD4")

@@

##Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.

```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE, fig.width=20}
pma<-filter(CD4, Stim=="PMA")
wcl<-filter(CD4, Stim=="WCL")
pep<-filter(CD4, Stim=="PEP")
triple_boolean_plot(pma)
triple_boolean_plot(wcl)
triple_boolean_plot(pep)
```

##T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups. 
```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
colnames<-c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq")
### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "WCL", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="WCL",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### Peptide
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PEP", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PEP",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```
\newpage

##Individual Cytokines
These are the frequencies of cells producing individual cytokines without considering whether or not they are single positive, double positive etc. So any cell producing IL-4 is included in the IL-4 analysis.

@@
